\mytitle

### 2016-08-18 pdocs/thesis/article-section-3.tex
article section 3 tuning - explain ep2 behaviour

Single simulation step is done using following substeps. 
In this work, changes are made only to constraing solving and update actions.
\begin{enumerate}
\item Apply gravity to each non static body. 
This step just makes programming easier as otherwise each program should add forces due to gravity in each step. 
In this case, $\vec{f}_{ext}$ gets added by $\lbrace{0,-60000,0}\rbrace$.
\item Predict unconstrained motion for each non static body. Prediction is done based on current linear and angular velocities of the body.
\item Predict contacts. In this phase, continuous collision detection is done based on simplified objects. 
Each rigid body is represented by sphere and contact prediction is done if body moves more than given threshold value during simulation step. Continuous collision detection is configured for each body and as default it is not active. 
Typical scenario that requires continuous collision detection is fast moving bodies that would otherwise go through walls.
\item Perform discrete collision detection. All overlapping body pairs are processed and manifoldPoints are created for
each detected contact at end of simulation step. 
In unconstrained case, contact is detected after t=0.55 s if continuous collision detection is not used and manifolds distance gets negative value (-0.058 m).
\item Calculate simulation islands(groups). Bodies that are near each other based on contact prediction or discrete collision detection
 or connected with constraints are grouped in same group.
\item Solve constraints. Both contact and other constraints are processed in this step.
\item Integrate transforms using velocities modified in previous step.
\item Update actions (callbacks) are called. 
In elastic-plastic case, plasticity is summed and equilibrium point of elastic part is updated if maximum force or moment is exceeded.
\item Activation state of bodies is updated. 
To avoid extra calculation bodies are as default put to sleeping state if linear and angular velocities of body are less than
given threshold values (default 0.8 and 1.0) more than set limit (2.0).
\end{enumerate}

@@

\begin{description}
\item[velError] is calculated using velocities and external impulses of connected bodies.  
 \begin{description}
\item[In constraint cases,] it is dominant contributor for constraint's impulse.
\item[In contact case,] main contributor is bodies relative speed at point of contact. 
Bodies are not allowed to penetrate each other.
\end{description}
\item[posError] is calculated by constraint. It is significant factor in designing stable constraints.
 \begin{description}
 \item[In fixed case,] value is about 12 times actual position error. Factor 12 is based on time step (60) 
 and default value of error reduction parameter (erp) which has value of 0.2 in this context.
 \item[In elastic plastic case,]  value is set to zero if impulse would be larger than maximum impulse or
spring simulation cannot be done in stable way.
 \item[In contact case,] value is zero if there is no penetration. For penetration cases it is 
$\frac{-penetration\, erp}{timeStep}$
 \end{description}
\item[rhs (c)] is calculated by velError jInv + posError jInv
\item[jInv] is calculated using masses and inertias of connected bodies and constraint geometry. 
 \begin{description}
\item[In constraint cases,] it is mass of body (6000).
\item[In contact case,] it varies below mass of body.
\end{description}
\item[Impulse] is impulse applied to body during timestep.
 \begin{description}
\item[In constraint cases,] it is obtained from btJointFeedback structure.
\item[In contact case,] it is obtained by summing applied impulses from active manifolds.
\end{description}
\item[erp] Error reduction parameter (0...1) is used to handle numerical issues e.q. object drifting. 
Setting erp to 1 would in theory eliminate error but in practice value of 0.8 is used in most cases.
\end{description}

Actual values for unconstrained case without continuous collision detection
are shown in \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s when
{\it velError} is 5.67 (5.5+0.167) and
{\it posError} is 2.8 (0.058*0.8/0.0167). Impulse is 34000 and contact force is thus about 2 MN (34000/0.0167).
After few steps location and position stabilize although internally calculation is needed for each time step
until body is deactivated.

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Parameter} & 
{\bf Description} 
lowerLimit &  
minimum allowed translation or rotation
 \\  \hline
upperLimit &  
maximum allowed translation or rotation
 \\  \hline
springStiffness & elastic spring stiffness
 \\  \hline
enableSpring & defines if spring is active
 \\  \hline
springStiffnessLimited & should elastic behaviour be tuned to avoid instability
 \\  \hline
equilibriumPoint & should elastic behaviour be tuned to avoid instability
 \\  \hline
currentLimit & describes state of constraint - 0: not limited,
 3: loLimit=hiLimit, 4: current value is between loLimit and HiLimit
 \\ \hline

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\it velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017 & -0.0028 &        &    &          & -0.17 & 0 & 0 \\  \hline
0.033 & -0.0046 &-0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
0.050 & -0.0056 &-0.28 & 0 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
0.067 & -0.0056 &-0.22 & 0 & -1340 &  -0.001&  1330 & 0.004\\  \hline
0.083... & -0.0056  & -0.17 & 0 & -1000 &  0.0&  1000 & 0.004\\  \hline
\end {tabular}
\end{center}
\caption {Constraint parameter values for elastic-plastic constraint} \label{tab:epBlockValues} 
\end {table}

Six-dof-spring constraint 2 has feature to avoid unstability by automatically softening constraint
spring if needed.
If this feature is active for elastic-plastic constraint 2 constraint it behaves as shown in table
 \ref{tab:ep2BlockValues}.

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017... &  0 & -0.17  & 2.34 & 13000 & 0       & 1000 & 0 \\  \hline
\end {tabular}
\end{center}
\caption {Constraint parameter values for elastic-plastic constraint 2} \label{tab:ep2BlockValues} 
\end {table}

>>>>

Single simulation step is done using following substeps. 
In this work, changes are made only to constraint setup and update actions.
\begin{enumerate}
\item Apply gravity to each non static body. 
This step just makes programming easier as otherwise each program should add forces due to gravity in each step. 
In this case, $\vec{f}_{ext}$ in  \ref{eq:eomV2} gets added by $\lbrace{0,-60000,0}\rbrace$.
\item Predict unconstrained motion for each non static body. Prediction is done based on current linear and angular velocities of the body.
\item Predict contacts. In this phase, continuous collision detection is done based on simplified objects. 
Each rigid body is represented by sphere and contact prediction is done if body moves more than given threshold value during simulation step. Continuous collision detection is configured for each body and as default it is not active. 
Typical scenario that requires continuous collision detection is fast moving bodies that would otherwise go through walls.
\item Perform discrete collision detection. All overlapping body pairs are processed and manifoldPoints are created for
each detected contact at end of simulation step. 
In unconstrained case, contact is detected after t=0.55 s if continuous collision detection is not used and manifolds distance gets negative value (-0.058 m).
\item Calculate simulation islands(groups). Bodies that are near each other based on contact prediction or discrete collision detection
 or connected with constraints are grouped in same group.
\item Solve constraints. Both contact and other constraints are processed in this step. 
 This step is subdivided to setup, iterations and finish phases. 
In setup phase constraints are queried for positional errors for calculation of c in
equation \ref{eq:constraintEquation} and maximum and minimum impulses in equations \ref{eq:lambdaLow} and 
\ref{eq:lambdaHigh}.
In finish phase constraint forces are calculated if requested and velocities of bodies are updated.
\item Integrate transforms using velocities modified in previous step.
\item Update actions (callbacks) are called. 
In elastic-plastic case, plasticity is summed and equilibrium point of elastic part is updated if maximum force or moment is exceeded.
\item Activation state of bodies is updated. 
To avoid extra calculation bodies are as default put to sleeping state if linear and angular velocities of body are less than
given threshold values (default 0.8 and 1.0) longer than set time limit (2.0).
\end{enumerate}

@@

\begin{description}
\item[velError] is calculated using velocities and external impulses of connected bodies.  
 \begin{description}
\item[In constraint cases,] it is dominant contributor for constraint's impulse.
\item[In contact case,] main contributor is bodies relative speed at point of contact. 
Bodies are not allowed to penetrate each other.
\end{description}
\item[posError] is calculated by constraint. It is significant factor in designing stable constraints.
 \begin{description}
 \item[In fixed case,] value is about 12 times actual position error. Factor 12 is based on time step (60) 
 and default value of error reduction parameter (erp) which has value of 0.2 in this context.
 \item[In elastic plastic case,]  value is set to zero if impulse would be larger than maximum impulse or
spring simulation cannot be done in stable way.
 \item[In contact case,] value is zero if there is no penetration. For penetration cases it is 
$\frac{-penetration\, erp}{timeStep}$
 \end{description}
\item[rhs (c)] is calculated by velError jInv + posError jInv
\item[jInv] is calculated using masses and inertias of connected bodies and constraint geometry. 
 \begin{description}
\item[In constraint cases,] it is mass of body (6000).
\item[In contact case,] it varies below mass of body.
\end{description}
\item[Impulse] is impulse applied to body during timestep.
 \begin{description}
\item[In constraint cases,] it is obtained from btJointFeedback structure.
\item[In contact case,] it is obtained by summing applied impulses from active manifolds.
\end{description}
\item[erp] Error reduction parameter (0...1) is used to handle numerical issues e.q. object drifting. 
Setting erp to 1 would in theory eliminate error in one step but in practice value of 0.2 - 0.8 is used in most cases.
\end{description}

Actual values for unconstrained case without continuous collision detection
are shown in table \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s when
{\it velError} is 5.67 (5.5+0.167) and
{\it posError} is 2.8 (0.058*0.8/0.0167). Impulse is 34000 and contact force is thus about 2 MN (34000/0.0167).
After few steps location and position stabilize although internally calculation is needed for each time step
until body is deactivated.

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|p{12cm}|}
\hline
{\bf Parameter} & 
{\bf Description} 
\\ \hline
lowerLimit &  
minimum allowed translation or rotation
 \\  \hline
upperLimit &  
maximum allowed translation or rotation
 \\  \hline
springStiffness & elastic spring stiffness
 \\  \hline
enableSpring & defines if spring is active
 \\  \hline
springStiffnessLimited & should elastic behaviour be tuned to avoid instability
 \\  \hline
equilibriumPoint & should elastic behaviour be tuned to avoid instability
 \\  \hline
currentLimit & describes state of constraint \newline
 0: not limited \newline
 3: loLimit=hiLimit \newline 4: current value is between loLimit and HiLimit
 \\ \hline

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\it velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017 & -0.0028 &-0.17 & 0 & -1000 & -0.17 & 0 & 0 \\  \hline
0.033 & -0.0046 &-0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
0.050 & -0.0056 &-0.28 & 0 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
0.067 & -0.0056 &-0.22 & 0 & -1340 &  -0.001&  1330 & 0.004\\  \hline
0.083... & -0.0056  & -0.17 & 0 & -1000 &  0.0&  1000 & 0.004\\  \hline
\end {tabular}
\end{center}
\caption {Constraint parameter values for elastic-plastic constraint} \label{tab:epBlockValues} 
\end {table}

Six-dof-spring constraint 2 has optional feature to avoid unstability by automatically softening constraint
spring if needed. Feature is activated if current timeStep is too large for spring-mass system simulation.
Equation \ref{eq:frequencyLimited} is used for decision, where $k$ is elastic spring stiffness between bodies,
$m_{min}$ is smaller of connected masses or inertias and $dt$ is integration timestep.
\begin{equation} \label{eq:frequencyLimited}
\sqrt{ k /  m_{min}} dt > 0.25 
\end{equation}
If this feature is active for elastic-plastic constraint 2 constraint  
it does not report positional error but it allows full plastic capacity to be used for correcting
velocity based error instead of force provided by spring.
In this scenario object behaves as shown in table  \ref{tab:ep2BlockValues} i.e. it does not move at all.

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017... &  0 & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
\end {tabular}
\end{center}
\caption {Constraint parameter values for elastic-plastic constraint 2} \label{tab:ep2BlockValues} 
\end {table}
